---
title: "Study of the foreign-born population in the United States"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---
\
\
\


 **Here we conduct  a brief study of the foreign-born population in the United States based on the American Community Survey (ACS) published by the Census Bureau. We use the person-level data published for 2014, which includes 3,134,610 individuals. We focus on the foreign-born population, identified using the "NATIVITY" variable in the data, which consists of 360,677 individuals.** 
\
\


```{r libraries, message=FALSE, include=FALSE}
# Load libraries:
library(dplyr)
library(data.table)
library(DT)
library(ggplot2)
library(plotly)
library(scatterD3)
```


```{r load data, message=FALSE, include=FALSE}
# Load data and select variables:
 #selected.variables = c("ST","POBP","ENG","SCHL", "NATIVITY", "YOEP", "AGEP", "PWGTP")
 #hus.a.data = fread("../data/ss14pusa.csv", select = selected.variables)
 #hus.b.data = fread("../data/ss14pusb.csv", select = selected.variables)
 #db = rbind(hus.a.data, hus.b.data)
 #rm(hus.a.data, hus.b.data)
 #save(db, file = "../data/db.RData")
  load("../data/db.RData")
```



```{r, message=FALSE, include=FALSE}
# Get country names and state abbreviations:

countries = fread("../data/countrynames.csv")
db = db %>% 
    left_join(., countries, by = c("POBP" = "code")) %>%
    rename(COB_name = name)

states = fread("../data/statenames.csv")
db = db %>% 
    left_join(., states, by = c("ST" = "code")) %>% 
    rename(ST_abbr = abbr) %>% 
    select(-c(name))

statenames = states$name

# remove DC and Puerto Rico rows:
db = filter(db, !ST_abbr %in% c('DC','PR'))
statenames = statenames[!statenames %in% c('District of Columbia','Puerto Rico')]
```


```{r, message=FALSE, include=FALSE}
# Find top countries of immigration:
top_n = 20 # the top how many countries of immigration
top_countries = db %>% 
                filter(NATIVITY==2) %>%
                group_by(COB_name) %>%
                summarise(sum(PWGTP)) %>%
                arrange(desc(`sum(PWGTP)`)) %>%
                head(top_n)
```


### Top `r top_n` countries of origin for immigrants in the United States:

| Rank |   |Country | Immigrant Population|
|------|---|--------|---------------------|
`r paste('|', 1:top_n, '|', paste('![](../data/Flags/24/',top_countries[[1]],'.png)', sep=''), '|', top_countries[[1]], '|', round(top_countries[[2]]/1e6, 3), 'M', '|', '\n', collapse='')`
###### (Icons: <http://www.icondrawer.com>)


```{r, message=FALSE, include=FALSE}
# Estimate percent immigrant in state poulation:

pop_total = db %>% group_by(ST_abbr) %>% summarise(sum(PWGTP)) %>% rename(tot = `sum(PWGTP)`)
pct_imm_by_st = db %>% 
              filter(NATIVITY==2) %>% 
              group_by(ST_abbr) %>% 
              summarise(sum(PWGTP)) %>% 
              mutate(percent_imm = round(100*`sum(PWGTP)`/pop_total$tot, 1)) %>%
              select(-`sum(PWGTP)`)

```



```{r, message=FALSE, include=FALSE}
# For each US state, get top countries of immigration:
top_n_st = 5 # the top how many countries of immigration to display
top_countries = list()
for(s in unique(db$ST_abbr)){
  state_rows = filter(db, ST_abbr==s & NATIVITY==2) # select only immigrants in state s
  top_countries[[s]] = state_rows %>%
                      group_by(COB_name) %>%
                      summarise(sum(PWGTP)) %>%
                      mutate(pct = round(100* `sum(PWGTP)` /sum(state_rows$PWGTP), 1)) %>%
                      select(-`sum(PWGTP)`) %>%
                      arrange(desc(pct)) %>%
                      head(top_n_st)
}
```
\
\


### Geographic distribution of immigrant groups:
\
\
```{r, message=FALSE, echo=FALSE}
# Display map:

nstates = length(top_countries)
hover = character(nstates)

# prepare list for display of top countries of immigration:
for(i in 1:nstates){
  hover[i] = paste('<b>', statenames[i], ' ', pct_imm_by_st$percent_imm[i], '%</b>, of which:<br>',
                    paste('  ',top_countries[[i]]$COB_name, ' ', top_countries[[i]]$pct, '% <br>', collapse=""),
                    collapse="")
}

# include in data frame along with other vars for plotting
state_imm = data.frame(hover = hover, code = unique(db$ST_abbr), stringsAsFactors=FALSE)
state_imm = left_join(state_imm, pct_imm_by_st, by = c('code'='ST_abbr'))


# give state boundaries a white border
l = list(color = toRGB("white"), width = 2)

# specify some map projection/options
g = list(scope = 'usa', projection = list(type = 'albers usa'), showlakes = TRUE, lakecolor = toRGB('white'))

plot_ly(state_imm, type = 'choropleth', z = percent_imm, 
        locationmode = 'USA-states', locations = code, 
        hoverinfo = 'text', text = hover, 
        color = percent_imm, colors = 'Oranges',
        marker = list(line = l), 
        colorbar = list(len = 0.75, thickness = 20, title = "Percent foreign-born")) %>%
  
  layout(geo = g, 
         title = 'Percent foreign-born per state and top countries of origin <br> (hover cursor over map for breakdown)', 
         autosize = FALSE,
         width = 900,
         margin = list(t=25, l=0, r=0, d=0))

```


### Age, Year of entry and Age of arrival of different immigrant groups:

```{r, message=FALSE, include=FALSE}
# Compute weighed densities of Age, Year of Entry, and Age at Entry for different immigrant groups. Weighing is done using survey weights PWGTP

select_countries = c('Mexico', 'China', 'Cuba', 'India', 'Germany') # top_countries$COB_name
plot_data = filter(db, NATIVITY==2 & COB_name %in% select_countries)

ages = integer()
yoes = integer()
aoas = integer()
age_freqs = numeric()
yoe_freqs = numeric()
aoa_freqs = numeric()
countries = character()


for(country in select_countries){
  plot_data_c = filter(plot_data, COB_name==country)
  
  # weighed density of ages
  dens_c = with(plot_data_c, density(AGEP, weights=PWGTP/sum(PWGTP), bw=4, n=100, from=0, to=100))
  ages = c(ages, dens_c[[1]])
  age_freqs = c(age_freqs, dens_c[[2]])
  
  # weighed density of year of entry
  dens_c = with(plot_data_c, density(YOEP, weights=PWGTP/sum(PWGTP), bw=3, n=100, from=1920, to=2020))
  yoes = c(yoes, dens_c[[1]])
  yoe_freqs = c(yoe_freqs, dens_c[[2]])
  
  # weighed density age at entry
  dens_c = with(plot_data_c, density(AGEP-(2014-YOEP), weights=PWGTP/sum(PWGTP), bw=3, n=100, from=-10, to=100))
  aoas = c(aoas, dens_c[[1]])
  aoa_freqs = c(aoa_freqs, dens_c[[2]])
  
  countries = c(countries, rep(country,100))
}

# data frame for plotting:
density_data = data.frame(age = ages, yoe = yoes, aoa = aoas, age_freq = age_freqs, 
                          yoe_freq = yoe_freqs, aoa_freq = aoa_freqs, country = countries)
```


We show the age distributions of five important immigrant groups (although not the five largest): Mexico, China, Cuba, India and Germany. These densities take into consideration the survey weights assigned to each observation.

```{r, warning=FALSE, echo=FALSE, fig.height=4, fig.align='center'}

# age distributions:
ggplot(density_data, aes(age, age_freq, color=country)) + 
  geom_line() +
  theme_bw() + 
  labs(title='Age distribution per immigrant group', x='Age', y='Frequency', color='Country') +
  scale_x_continuous(breaks=seq(0,100,10))
```
\
We notice that German immigrants tend to be significantly older than other groups, with a most likely age of about 70 years old. Cuban foreign-born residents are the second oldest group, with a typical age of about 50. In constrast, the bulk of Mexican and Indian immigrants tends to be in their thirties. Interestingly, the Chinese foreign-born population appears to have a bimodal distribution, with most idividuals either in their mid-twenties or in their forties. \
These age differences might be explained by the year in which each group entered the United States. We look at this next.
\
\

 
```{r, warning=FALSE, echo=FALSE, fig.height=4, fig.align='center'}
# year of entry distributions:
ggplot(density_data, aes(yoe, yoe_freq, color=country)) + 
  geom_line() +
  theme_bw() + 
  labs(title='Frequency of year of entry per immigrant group', x='Year of entry', y='Frequency', color='Country') +
  xlim(c(1920,2010)) +
  annotate("text", x=1957, y=0.027, label='Post-WWII') +
  annotate("text", x=1967, y=0.018, label='Cuban Rev.') 
```
\
As suggested by the age distributions, we see that the large majority of the German group arrived in the coutry the earliest, in the two decades following World War II. The Cuban group followed, most likely pushed to exile by the revolution of 1953-59. We also see that the Mexican wave of immigration seems to have peaked around 2000, whereas Chinese and Indian immigration appears to not yet have reached its peak.\
Finally, we look at the distribution of the age of arrival, computed by subtracting the years since arrival (2014 - year-of-entry) to the current age.
\
\


```{r, warning=FALSE, echo=FALSE, fig.height=4, fig.align='center'}
# age at entry distributions:
ggplot(density_data, aes(aoa, aoa_freq, color=country)) + 
  geom_line() +
  theme_bw() + 
  labs(title='Distribution of age of arrival per immigrant group', x='Age of arrival', y='Frequency', color='Country') +
  xlim(c(0,75)) 
 
```
\
We see a general trend in most groups to immigrate to the country in their twenties. There is also evidence of a smaller mode at 0-10 years of age for most distributions, which suggests that a lot of foreign-born residents entered the country as small children.


### Now, let's move on and take a look at the education level
\



\
###Let's take a look at the employment status of the migrations in the U.S.
Firstly to choose the varaibles we are intersted in: 

\
After comparing several similiar variables, we chose Class of Worker(COW),Employment Status Record(ESR) and Industry Field Record (INDP) as three attributes to evalute the employment situations of the immigrants in the U.S.
\
```{r, echo=FALSE}
#Load data and select variables:
selected.variables = c("ST","SCHL", "NATIVITY","PWGTP","INDP","SOCP","COW","ESR","POBP","POVPIP","WAGP","FINCP","HINCP","FFINCP","FHINCP","OIP")
hus.a.data = fread("~/Desktop/ss14pusa.csv", select = selected.variables)
hus.b.data = fread("~/Desktop/ss14pusb.csv", select = selected.variables)
db = rbind(hus.a.data, hus.b.data)
rm(hus.a.data, hus.b.data)

# Get country names and add to db:
countries = fread("~/Desktop/countrynames.csv")
db = db %>%
  left_join(., countries, by = c("POBP" = "code")) %>%
  mutate(COB_name = name) %>%
  select(-name)
```
  

\
Then choose the countries we are interested in:

```{r, echo=FALSE}
# Find top countries of immigration:
top_n = 20 # the top how many countries of immigration
top_countries = db %>% 
  filter(NATIVITY==2) %>%
  group_by(COB_name) %>%
  summarise(sum(PWGTP)) %>%
  arrange(desc(`sum(PWGTP)`)) %>%
  head(top_n)
  
```
\
 From the above picture, we can see that these are the top 20 countries migrant to the U.S. I picked Mexico, China, Cuba, Canada and Germany. As they are in South America, Asia, Middle America, North America and Europe, I think these five countries are representative. So I chose them as the our goals.
 
 ```{r, echo=FALSE}
select_countries = c('Mexico','China', 'Cuba', 'Canada', 'Germany') # top_countries$COB_name
plot_data = filter(db, NATIVITY==2 & COB_name %in% select_countries)

 ```
 ```{r, echo=FALSE}
 #bar chart about class of worker in the typical countries, use the variable COW
species = c(rep("Mexico",10),rep("China",10),rep("Cuba",10),rep("Canada",10),rep("Germany",10))
conditions = rep(c("for wages","none-profit","local government","state government","fedral government","Self-employed in own not incorporated","Self-employed in own incorporated","Working without pay(family business or farm)","unemployed","less than 16"),5)

count_country_class_num = matrix(nrow = 5,ncol = 10)
for(i in 1:5){
  for(j in 0:9){
    count_country_class_num[i,j]=dim(filter(plot_data,COW==j & COB_name==select_countries[i]))[1]
  }
}
less_than_16 = vector()
for (i in 1:5){
  less_than_16[i] = sum(is.na(plot_data$COW) & plot_data$COB_name==select_countries[i])
}
count_country_class_num[,10]=less_than_16

values = as.numeric(matrix(data = count_country_class_num,nrow = 1,byrow = FALSE))
data = data.frame(species,conditions,values)
p <- ggplot(data, aes(fill=conditions, y=values, x=species))
p + geom_bar(position="dodge", stat="identity") + labs(title='Class of worker', x='Countries', y='Values', fill='Class')
```
\
From the above Class of Workers graph, we may conclude that Mexico has significant high level of people who immigrate to the U.S working for wages. An interesting result also shows that Germany has the most immigrants to who work in the U.S.for self-employed in their own but not incorporated companies.

```{r, echo=FALSE}
#bar chart about the Employment status recode, using the variable ESR
under16 = vector()
for ( i in 1:5){
  under16[i]=sum(is.na(filter(plot_data,plot_data$COB_name==select_countries[i])$ESR))
  print(table(filter(plot_data,plot_data$COB_name==select_countries[i])$ESR))
}
values_esr = c(3574,52489,1241,3855,61,3,29786,1104,8909,288,531,20,0,7370,345,4853,93,382,9,0,4346,14,4118,149,201,10,0,3928,121,2489,74,128,14,0,3629)
species_esr= c(rep("Mexico",7),rep("China",7),rep("Cuba",7),rep("Canada",7),rep("Germany",7))
conditions_esr = rep(c("under 16","civilian employed and at work","civilian employed without work","unemployed","armed forces, at work","armed forces without work","not in labor force"),5)
data_esr = data.frame(species_esr,conditions_esr,values_esr)
p <- ggplot(data_esr, aes(fill=conditions_esr, y=values_esr, x=species_esr))
p + geom_bar(position="dodge", stat="identity") + labs(title='Employment Status Recode', x='Countries', y='Values', fill='Employment Status')

```

\
As we can see from the employment status recode graph, Mexican people surplus other countries of the total immigrants to the U.S. The civilian employed at work and not in labor force also take a large percentage. China ranks second and the other three countries all have the similar patterns, showing that people who are civilian employed and at work and not in labor force are the main status of the migrants in the U.S.

```{r, echo=FALSE}
#bar chart about Industry recode using the variable INDP
#classify industry through INDP fisrt three letters
AGR = c(170,180,190,270,280,290)
EXT = c(370,380,390,470,490)
UTL = c(570,580,590,670,680,690)
CON = c(770)
MFG = c(1070,1080,1090,1170,1180,1190,1270,1280,1290,1370,1390,1470,1480,1490,1570,1590,1670,1680,1690,1770,1790,1870,1880,1890,1990,2070,2090,2170,2180,2190,2270,2280,2290,2370,2380,2390,2470,2480,2490,2570
        ,2590,2670,2680,2690,2770,2780,2790,2870,2880,2890,2970,2980,2990,3070,3080,3095,3170,3180,3190,3365,3370,3380,3390,3470,3490,3570,3580,3590,3670,3680,3690,3770,3780,3790,3875,3895,3960,3970,3980,3990)
WHL = c(4070,4080,4090,4170,4180,4195,4265,4270,4280,4290,4370,4380,4390,4470,4480,4490,4560,4570,4580,4585,4590)
RET = c(4670,4680,4690,4770,4780,4795,4870,4880,4890,4970,4980,4990,5070,5080,5090,5170,5180,5190,5275,5280,5295,5370,5380,5390,5470,5480,5490,5570,5580,5590,5591,5592,5670,5680,5690,5790)
TRN = c(6070,6080,6090,6170,6180,6190,6270,6280,6290,6370,6380,6390)
INF = c(6470,6480,6490,6570,6590,6670,6672,6680,6690,6695,6770,6780)
FIN = c(6870,6880,6890,6970,6990,7070,7080,7170,7180,7190)
PRF = c(7270,7280,7290,7370,7380,7390,7460,7470,7480,7490,7570,7580,7590,7670,7680,7690,7770,7780,7790)
EDU = c(7860,7870,7880,7890)
MED = c(7970,7980,7990,8070,8080,8090,8170,8180,8190,8270,8290)
SCA = c(8370,8380,8390,8470)
ENT = c(8560,8570,8580,8590,8660,8670,8680,8690)
SRV = c(8770,8780,8790,8870,8880,8970,8980,8990,9070,9080,9090,9160,9170,9180,9190,9290)
ADM = c(9370,9380,9390,9470,9480,9490,9470,9480,9490,9570,9590)
MIL = c(9670,9680,9690,9770,9780,9790,9870)
UNEMPLOYED = 9920
foreignkey1 = c(AGR,EXT,UTL,CON,MFG,WHL,RET,TRN,INF,FIN,PRF,EDU,MED,SCA,ENT,SRV,ADM,MIL,UNEMPLOYED)
foreignkey2 = c(rep("AGR",length(AGR)),rep("EXT",length(EXT)),rep("UTL",length(UTL)),rep("CON",length(CON))
                ,rep("MFG",length(MFG)),rep("WHL",length(WHL)),rep("RET",length(RET)),rep("TRN",length(TRN))
                ,rep("INF",length(INF)),rep("FIN",length(FIN)),rep("PRF",length(PRF)),rep("EDU",length(EDU))
                ,rep("MED",length(MED)),rep("SCA",length(SCA)),rep("ENT",length(ENT)),rep("SRV",length(SRV))
                ,rep("ADM",length(ADM)),rep("MIL",length(MIL)),rep("UNEMPLOYED",length(UNEMPLOYED)))
foreignkeys = as.data.frame(cbind(foreignkey1,foreignkey2))
names(foreignkeys)[1] = "INDP"
plot_data1 = merge(plot_data, foreignkeys, by=c("INDP"))
Mexico = filter(plot_data1,COB_name=="Mexico")
China = filter(plot_data1,COB_name =="China")
Cuba = filter(plot_data1,COB_name == "Cuba")
Canada = filter(plot_data1,COB_name == "Canada")
Germany = filter(plot_data1,COB_name =="Germany")
Mexico_num = as.numeric(table(Mexico$foreignkey2))
China_num = as.numeric(table(China$foreignkey2))
Cuba_num = as.numeric(table(Cuba$foreignkey2))
Canada_num = as.numeric(table(Canada$foreignkey2))
Germany_num = as.numeric(table(Germany$foreignkey2))
num_mat = rbind(Mexico_num,China_num,Cuba_num,Canada_num,Germany_num)
less_than_16_INPD = vector()
for (i in 1:5){
  less_than_16_INPD[i]=sum(is.na(filter(plot_data,COB_name == select_countries[i])$INDP))
}
total_num_mat = cbind(less_than_16_INPD,num_mat)
total_num_val = matrix(t(total_num_mat),nrow = 1,byrow = TRUE)
species_INDP = c(rep("Mexico",20),rep("China",20),rep("Cuba",20),rep("Canada",20),rep("Germany",20))
conditions_INDP = rep(c("under 16","AGR","EXT","UTL","CON","MFG","WHL","RET","TRN","INF","FIN","PRF","EDU","MED","SCA","ENT","SRV","ADM","MIL","UNEMPLYED"),5)
data_INDP = data.frame(species_INDP,conditions_INDP,total_num_val)
p<-ggplot(data_INDP,aes(fill=conditions_INDP,y=total_num_val,x=species_INDP))
p+geom_bar(position = "dodge",stat = "identity")+labs(title="Industry fields record in different countries")
```
\
From this industry record graph, we know that the teenagers take the most percentage in all five countries. Mexicans also take a large percentage in finance, manufacturing as well as United Telecoms Limited industrial areas.

###Now let's have a look at the immigrations' income status.
```{r load data, message=FALSE, include=FALSE}
#analysis about incomes; use the variable WAGP(wage/salary of a year)
###Here is an interactive plot that will show 20 countries' immigrants' income
install.packages("plotly")
library(plotly)
income_plot = filter(db, NATIVITY==2 & COB_name==top_countries$COB_name)
income_plot$WAGP=income_plot$WAGP/10000
plot_ly(income_plot[1:2000,],x=COB_name,y=WAGP,mode = "markers",color = COB_name,size=WAGP)

``` 
 \
 We can also get each country's highest wage from this command
 ```{r load data, message=FALSE, include=FALSE}
 plot_ly(income_plot,color = COB_name,x= COB_name, y = WAGP)
```
\
From the above graph, we may conclude that, China, India, Korea and Canada have almost the same highest salaries. Honduras has the lowest highest wage among the top immigrants.

\
Or we can choose an easier way: just use their wage divided by 10000 to represent the relative differences between wages. And in this way can we make the interactive plot.

```{r load data, message=FALSE, include=FALSE}
#analysis about incomes; use the variable WAGP
install.packages("plotly")
library(plotly)
income_plot = filter(db, NATIVITY==2 & COB_name==top_countries$COB_name)
income_plot$WAGP=income_plot$WAGP/10000
plot_ly(income_plot[1:2000,],x=COB_name,y=WAGP,mode = "markers",color = COB_name,size=WAGP)

##highest salary
plot_ly(income_plot,color = COB_name,x= COB_name, y = WAGP)
```
\
When we want to know about the income status about the immigrants in the U.S, we chose the wage/salary income(WAGP) variable in the past year. We can divide them into nine class:
Class 1:0 (no payment)
Class 2: 0-20000 (0 is not included)
Class 3:20000-40000
Class 4: 40000-60000
Class 5: 60000-80000
Class 6: 80000-100000
Class 7: 100000-150000
Class 8: >150000
Class 9: no record

```{r load data, message=FALSE, include=FALSE}
#analysis about incomes; use the variable WAGP
install.packages("plotly")
library(plotly)
income_plot = filter(db, NATIVITY==2 & COB_name==top_countries$COB_name)
income_plot$WAGP=income_plot$WAGP/10000
plot_ly(income_plot[1:2000,],x=COB_name,y=WAGP,mode = "markers",color = COB_name,size=WAGP)

##highest salary
##classification
income_plot$WAGP=income_plot$WAGP*10000
income_plot$WAGP[income_plot$WAGP>0 & income_plot$WAGP<=20000]=1
income_plot$WAGP[income_plot$WAGP>20000 & income_plot$WAGP<=40000]=2
income_plot$WAGP[income_plot$WAGP>40000 & income_plot$WAGP<=60000]=3
income_plot$WAGP[income_plot$WAGP>60000 & income_plot$WAGP<=80000]=4
income_plot$WAGP[income_plot$WAGP>80000 & income_plot$WAGP<=100000]=5
income_plot$WAGP[income_plot$WAGP>100000 & income_plot$WAGP<=150000]=6
income_plot$WAGP[income_plot$WAGP>150000]=7
no_record = vector()
no_payment = vector()
m = matrix(nrow = 5,ncol = 7)
for (i in 1:5){
  no_record[i] = sum(is.na(filter(income_plot,COB_name==select_countries[i])$WAGP))
  matr =filter(income_plot,COB_name==select_countries[i])
  for(j in 1:7){
    m[i,j]=dim(filter(matr,WAGP==j))[1]
  }
  no_payment[i] = dim(filter(matr,WAGP==0))[1]
}
fin_num = cbind(no_payment,m,no_record)
vals = matrix(t(fin_num),ncol = 1)
spes = c(rep("Mexico",9),rep("China",9),rep("Cuba",9),rep("Canada",9),rep("Germany",9))
cons = rep(c("no payment","0-20000","20000-40000","40000-60000","60000-80000","80000-100000","100000-150000",">150000","no record"),5)
data_income = data.frame(spes,cons,vals)
p <- ggplot(data_income, aes(fill=cons, y=vals, x=spes))
p + geom_bar(stat = "identity", position="dodge") + labs(title="Income differences in different countries")
```
\
From above picture, we can say that in the typical five countries, immigrants who have no payment take the most percentage. And in Mexico and China, immigrants who are only paid low money (0-20000) ranks seond.





