---
title: "Idea: studying foreign-born people in the United States"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

#### Preliminary analysis: Plot a map with the percentage of foreign-born people in each state (as indicated by POBP variable >=100) as well as the top 5 countries of origin for each state (displayed by hovering mouse over map)
\
\


#### Load libraries:
```{r libraries, message=FALSE}
library(dplyr)
library(data.table)
library(DT)
library(ggplot2)
library(plotly)
```
\
\


#### Load data and select variables:
```{r load data, message=FALSE}
 selected.variables = c("PUMA", "ST", "POBP", "FPOBP", "PWGTP")
# "WAOB", "ENG","LANX", "SCHL", "YOEP", "DECADE", "HISP", "LANP", "MIGSP", "NATIVITY", "NOP", "WGTP", paste0("wgtp", 1:80))

 hus.a.data = fread("../data/ss14pusa.csv", select = selected.variables)
 hus.b.data = fread("../data/ss14pusb.csv", select = selected.variables)

 db = rbind(hus.a.data, hus.b.data)
 rm(hus.a.data, hus.b.data)

# load("../data/db.RData")

```
\
\


#### Get country names and state abbreviations:

```{r, message=FALSE}
countries = fread("../data/countrynames.csv")
db = left_join(db, countries, by = c("POBP" = "code"))
db = mutate(db, COB_name = name)
db = select(db, -name)

states = fread("../data/statenames.csv")
db = left_join(db, states, by = c("ST" = "code"))
db = mutate(db, ST_abbr = abbr)
db = select(db, -c(name, abbr))

statenames = states$name

# remove DC and Puerto Rico rows:
db = filter(db, !ST_abbr %in% c('DC','PR'))
statenames = statenames[!statenames %in% c('District of Columbia','Puerto Rico')]
```
\
\


#### Estimate percent immigrant in state poulation:
```{r, message=FALSE}
pop_total = db %>% group_by(ST_abbr) %>% summarise(sum(PWGTP)) %>% rename(tot = `sum(PWGTP)`)
pct_imm_by_st = db %>% 
              filter(POBP>=100) %>% 
              group_by(ST_abbr) %>% 
              summarise(sum(PWGTP)) %>% 
              mutate(percent_imm = round(100*`sum(PWGTP)`/pop_total$tot, 1)) %>%
              select(-`sum(PWGTP)`)

```
\
\


#### For each US state, get top countries of immigration:

```{r, message=FALSE}
top_n = 5 # the top how many countries of immigration to display
top_countries = list()
for(s in unique(db$ST_abbr)){
  state_rows = filter(db, ST_abbr==s & POBP>=100) # select only immigrants in state s
  top_countries[[s]] = state_rows %>%
                      group_by(COB_name) %>%
                      summarise(sum(PWGTP)) %>%
                      mutate(pct = round(100* `sum(PWGTP)` /sum(state_rows$PWGTP), 1)) %>%
                      select(-`sum(PWGTP)`) %>%
                      arrange(desc(pct)) %>%
                      head(top_n)
}
```
\
\


#### Display map:
```{r, message=FALSE}
nstates = length(top_countries)
hover = character(nstates)

# prepare list for display of top countries of immigration:
for(i in 1:nstates){
  hover[i] = paste('<b>', statenames[i], ' ', pct_imm_by_st$percent_imm[i], '%</b>, of which:<br>',
                    paste('  ',top_countries[[i]]$COB_name, ' ', top_countries[[i]]$pct, '% <br>', collapse=""),
                    collapse="")
}

# include in data frame along with other vars for plotting
state_imm = data.frame(hover = hover, code = unique(db$ST_abbr), stringsAsFactors=FALSE)
state_imm = left_join(state_imm, pct_imm_by_st, by = c('code'='ST_abbr'))


# give state boundaries a white border
l = list(color = toRGB("white"), width = 2)

# specify some map projection/options
g = list(scope = 'usa', projection = list(type = 'albers usa'), showlakes = TRUE, lakecolor = toRGB('white'))

plot_ly(state_imm, type = 'choropleth', z = percent_imm, 
        locationmode = 'USA-states', locations = code, 
        hoverinfo = 'text', text = hover, 
        color = percent_imm, colors = 'Oranges',
        marker = list(line = l), 
        colorbar = list(len = 0.75, thickness = 20, title = "Percent foreign-born")) %>%
  
  layout(geo = g, 
         title = 'Percent foreign-born per state and top countries of origin <br> (hover cursor over map for breakdown)', 
         autosize = FALSE,
         width = 900,
         margin = list(t=25, l=0, r=0, d=0))

```



